---
title: "COVID-19 Report in the United States"
author: "Holin Chen"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output: 
  html_document:
    theme: yeti
    hightlight: tango
    code_folding: show
    df_print: paged
    fig_width: 7
    fig_height: 6
    fig_caption: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Exploring COVID-19 Data in the United States

## Import data

Read the raw data from an online github respository (https://github.com/nytimes/covid-19-data). It daily updates the latest real-time data of COVID-19 cases and deaths in the USA  by compiling data from the New York Time.

```{r read data}
#install packages
library(RCurl)
library(dplyr)
library(ggplot2)
library(shadowtext)
library(plotly)

a <- read.csv(text=getURL("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"),header=T)
```

## Data cleaning and manipulation
Set the baseline as 100 cases for each state. Set time variable as the number of days since the state reached the 100th cases
```{r data manipulation}

aa <- a %>%
  as_tibble %>%
  filter(cases > 100) %>%
  group_by(state) %>%
  mutate(days_since_100 = as.numeric(as.Date(date) - min(as.Date(date)))) %>%
  ungroup

#Set the points of case numbers on y-axis
breaks = c(100,200,500,1000,2000,5000,10000,20000,50000,100000,200000,500000,1000000)
```

## Create the cumulative case curve ny state
```{r plot}

p <- ggplot(aa, aes(days_since_100, cases, color = state)) +
  geom_line(size = 0.8) +
  geom_point(pch = 21, size = 1) +
  scale_y_log10(expand = expansion(add = c(0,0.1)),
                breaks = breaks, labels = breaks) +
  scale_x_continuous(expand = expansion(add = c(0,1))) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none",
    plot.margin = margin(3,15,3,3,"mm")
  ) +
  coord_cartesian(clip = "off") +
  labs(x = "Number of days since 100th case", y= "",
       subtitle = "Total number of COVID-19 cases in USA by state") +
  geom_shadowtext(aes(label = paste0("", state)), hjust=0, vjust=0,
                         data = .%>% group_by(state) %>% top_n(1, days_since_100),
                         bg.color = "white")

p

```

## Create the interactive version of plot
```{r interactive}
fig <- ggplotly(p)
fig
```

## Explore data and create table

Now I want to know the case fatality rate (CFR) by calculating deaths divided by cases in each state. I create this new variable. Then I want to see which state or area have the maximum, median, and minimum cases among the US. I extract the rows of this data include the latest case and death counts, and CFR. Finally I put them into a table.
```{r table}
#create the new data with the latest case and death counts, and CFR
s <- a %>% 
  group_by(state) %>%
  summarise(latest_cases = max(cases, na.rm=T),
            latest_deaths = max(deaths, na.rm=T),
            latest_CFR = paste0(round(latest_deaths*100/latest_cases, digits=2),'%'))

#Put the 3 rows into a table
summary_table <- 
  s[c(which.max(s$latest_cases),
      which(s$latest_cases==median(s$latest_cases)),
      which.min(s$latest_cases)), ]

summary_table$cases_level <- c('max cases','median cases','min cases')
summary_table <- summary_table[,c(1,5,2:4)]

knitr::kable(summary_table, row.names(c('max cases','median cases','min cases')),align = 'c')
```
`r s$state[which.max(s$latest_cases)]` has the most cumulative cases with `r max(s$latest_cases)` cases so far. It's case fatality rate is `r s$latest_CFR[which.max(s$latest_cases)]`. `r s$state[which.min(s$latest_cases)]` has the least cumulative cases with `r min(s$latest_cases)` cases so far. It's case fatality rate is `r s$latest_CFR[which.min(s$latest_cases)]`.

<!--
I will use the data from this github respository for my final project in this class, which is inconsistent with what I said in homework 1 as my initial thought, casue my thesis progress is kinda slow now and I decide to change my idea to use some public data instead. Also I think using data direclty in github should make my work more reproducible!
-->











